---
title: "Software Tutorial: Verification and post-processing"
author: Alexander Jordan, Sebastian Lerch
date: ScienceFore Summer School, October 2017
output:
  beamer_presentation:
    highlight: default # “default”, “tango”, “pygments”, “kate”, “monochrome”, “espresso”, “zenburn”, and “haddock” 
---

```{r, eval = TRUE, echo = FALSE}
user <- "sl"  # "sl" / "aj"
```

## Objective and plan

In this software tutorial, we will evaluate and statistically post-process ensemble forecasts of wind speed. \pause Outline:

- explore data set of wind speed forecasts and observations
- evaluation of ensemble forecasts
- design of a simple post-processing model
- evaluation of post-processed forecasts

\bigbreak\pause
\textit{Please feel free to code along with us, possibly in groups. Interrupt us at any time if you have any questions.}

\bigbreak
The code will be shared on \url{https://github.com/slerch/sfss}

```{r, eval = TRUE, echo = FALSE}
if(user == "sl"){
  data_dir <- "/home/sebastian/Dropbox/ERC SummerSchool/GitHub_repo/postproc_example/data/new/"
} else if(user == "aj"){
  data_dir <- 123
}
load(paste0(data_dir, "HDwind.Rdata"))
```


## Data set \texttt{HDwind.Rdata}

The data set has been distributed via email. It contains

- ECMWF ensemble forecasts (50 members) of wind speed 
    - for a model grid point close to Heidelberg
    - with a forecast horizon of 60 hours (2.5 days)
    - valid at 12 UTC
    - from January 2015 - December 2016
- corresponding re-analysis values
- dates at which the forecasts and re-analysis values are valid

\pause
\small
All data have been downloaded from the TIGGE archive, see \url{https://github.com/slerch/sfss/tree/master/postproc_example/data_preprocessing} if you are interested


## Now, let's load and explore the data set

```{r, echo = FALSE, eval = TRUE}
library(scoringRules)
crpsv_ens <- crps_sample(y = obs, dat = ensfc)
# summary(crpsv_ens)
```

## Our post-processing model

Let's fit a simple EMOS model. As a basic example, we will start by fitting a normal model of the form
\[
Y | X_1,\dots,X_m \sim \mathcal{N}(\mu, \sigma^2),
\]
\pause
where the location parameter $\mu$ is a linear function of the ensemble mean,
\[
\mu = a + b\bar X,
\]
\pause 
and the scale parameter $\sigma$ is assumed to be constant,
\[
\sigma^2 = c.
\]
Plan: Use data from 2015 to estimate model for 2016.

## Steps in implementing our post-processing model

1. define objective function (mean CRPS as a function of the EMOS parameters $a,b,c$)
2. determine optimal parameters over the training set by numerically minimizing the objective function
3. compute out of sample parameter values in evaluation period
4. evaluate model


## Exercise

Build your own post-processing model. Try to outperform our simple benchmark model for data from 2016.
\bigbreak
\pause
Some hints for potential improvements:

- other information from the ensemble?
- more suitable parametric families for wind speed? (hint: `pnorm(0, n_mu, n_sigma)`)
- alternative ways to select the training period?
